---
title: "P8105 Homework 3"
author: "Diana Hernandez"
date: "2023-10-14"
output: github_document
---

# load tidyverse and setting options

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1: Instacart

```{r}
library(p8105.datasets)
data("instacart")
```

There are a total of `r length(instacart)` variables. There are `r nrow(instacart)` rows and `r ncol(instacart)` columns. Examples of key variables include order ID, product ID, user ID, product name, aisle ID, aisle number, etc. 

```{r}
instacart_df =
  instacart |>
  group_by(aisle) |>
  summarize(n_obs = n())
```

There are 134 aisles. The aisles that the most items are ordered from are: fresh vegetables, fresh fruits, and packaged vegetable fruits. 







