---
title: "P8105 Homework 3"
author: "Diana Hernandez"
date: "2023-10-14"
output: github_document
---

# load tidyverse, setting options, and p8105.datasets

```{r}
library(tidyverse)

library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1: Instacart

```{r}
data("instacart")
```

There are a total of `r length(instacart)` variables. There are `r nrow(instacart)` rows and `r ncol(instacart)` columns. Examples of key variables include order ID, product ID, user ID, product name, aisle ID, aisle number, etc. 

```{r}
instacart_df = 
  instacart
  
instacart_df |>
  group_by(aisle) |>
  summarize(n_obs = n())
```

There are 134 aisles. The aisles that the most items are ordered from are: fresh vegetables, fresh fruits, and packaged vegetable fruits. 

```{r}
instacart_plot = 
instacart_df |>
  filter(n() >= 10000)
  ggplot(instacart_df, aes(x = aisle_id)) + 
  geom_bar() +
  labs(
    title = "Orders by aisle",
    x = "Aisles by ID",
    y = "Number of orders")
```

```{r}
instacart_table =
instacart_df |> 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(product_name, aisle) |> 
  summarize(count = n()) 
```

# Problem 2: BRFSS

```{r}
data("brfss_smart2010")
```

# Data cleaning

```{r}
brfss_smart2010_df = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  select(-location_id) |>
  filter(topic == "Overall Health") |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))
  ) |>
  arrange(response)
```

There are a total of `r length(brfss_smart2010_df)` variables. There are `r nrow(brfss_smart2010_df)` rows and `r ncol(brfss_smart2010_df)` columns.



